---
title: "Progress Memo 1"
subtitle: |
  | Final Project 
  | Data Science 1 with R (STAT 301-1)
author: "Sonia Bhatia"
date: today

format:
  html:
    toc: true
    embed-resources: true
    
execute:
  echo: false
  warning: false

from: markdown+emoji 
---

::: {.callout-important}
## Memo format

Like all other submissions for this course, we expect the memo to be neat and well organized.

You should make this document your own. This is a template. To help with this we have added a date to the yaml above (good whenever writing memos). You should remove this block when making this document your own.

:::

::: {.callout-tip icon=false}

## Github Repo Link

To link to your github **repo**sitory, appropriately edit the example link below. Meaning replace `https://your-github-repo-url` with your github repo url. Suggest verifying the link works before submitting.

[https://github.com/stat301-1-2023-fall/final-project-1-soniab101](https://github.com/stat301-1-2023-fall/final-project-1-soniab101)

:::

::: {.callout-warning}
GitHub (free account) cannot store large files. If you try to commit and push a large dataset you will have an ERROR! Any file over 100 MB (100,000 KB) needs to be added to the `.gitignore` file BEFORE committing.

**We may need to do that for files in your final project!**
:::

## Data source
```{r}
library(tidyverse)
# loaded in the dataset using import data 
```
I got this dataset from kaggle using this link [https://www.kaggle.com/datasets/abecklas/fifa-world-cup?select=WorldCupMatches.csv]. I anticipate merging the World Cup Matches and World Cups datasets, since I think they both contain unique interesting data that will be good to analyze as a whole. I will merge them by the 'year' column.  


## Why this data

This dataset is interesting to me since I enjoy watching soccer matches, especially the world cup. I there are many conclusions that can be drawn from this dataset, and many interesting questions that can be explored. I am interested in seeing if there is any correlation between a team playing a match in their home country and whether that gives them an advantage in winning or not. I am also interested in the differences between viewership of matches.

## Data quality & complexity check

I was able to properly read in my data using import datasts. I have 20 variables in the World Cup Matches dataset, 10 of which are categorical and 10 are numerical. I have 10 vars in the WorldCups dataset, 5 of which are categorical and 5 are numerical. I am not running into missingness issues with the datasets. 

```{r}
#| eval: false
WorldCupMatches <- read_csv("data/WorldCupMatches.csv")
WorldCups<- read_csv("data/WorldCups.csv")
```


## Potential data issues

One potential difficulty is dealing with the datetime variable, as it is coded in a unique way. Another issue may be the fact that there are so many entries for each match id number.

## Misc

I may need to investigate the best way to merge these two datasets together, whether it be by the common year column or some other way.
