---
title: "Progress Memo 2"
subtitle: |
  | Final Project 
  | Data Science 1 with R (STAT 301-1)
author: "Sonia Bhatia"
date: today

format:
  html:
    toc: true
    embed-resources: true
    
execute:
  echo: false
  warning: false

from: markdown+emoji 
---



::: {.callout-tip icon=false}

## Github Repo Link

My github link is below.

[https://github.com/stat301-1-2023-fall/final-project-1-soniab101](https://github.com/stat301-1-2023-fall/final-project-1-soniab101)

:::

##Load in Data and Packages
```{r}
library(tidyverse)
library(skimr)
WorldCupMatches <- read_csv("data/WorldCupMatches.csv")
WorldCupPlayers <- read_csv("data/WorldCupPlayers.csv")
WorldCups <- read_csv("data/WorldCups.csv")

```

# summary of the datasets

# Investigating the data:
```{r}
summary(WorldCups)
summary(WorldCupMatches)

skim(WorldCups)
skim(WorldCupMatches)
# once you get everything from summary, go into each var and investigate it
# look at the distribution of scores

# Clean the team name Germany,
# when merging, could delete some cols if they aren't useful
# look into the column that has the NA values, could potentially delete the observations with NA values or even the entire column
# could change the NA values to the average so that they don't impact the analysis


WorldCups  |>  summarize(n=n(),.by=Winner)

```

# Cleaning the Data
I noticed that the team name Germany was coded both as Germany and also Germany FR. After investigating the meaning of this, I realized that Germany FR was West Germany before the country was reunified. I decided to recode Germany FR as Germany, since that is how the country is considered today,

# Analysis 1
Performance trend in top 4 teams that have won the most world cup matches over time

explore the data, missingness, univariate, bivariate, multivariate analysis

```{r}

```


# Example join:
```{r}
# Example data frames
world_cup_matches <- data.frame(
  Year = c(2018, 2018, 2014, 2014),
  Match = c("Final", "Semifinal", "Final", "Semifinal"),
  Team = c("France", "Croatia", "Germany", "Argentina"),
  Score = c(4, 2, 1, 0)
)

world_cups <- data.frame(
  Year = c(2018, 2014),
  Host = c("Russia", "Brazil"),
  Winner = c("France", "Germany")
)

# Perform a left join based on the "Year" column
merged_data <- left_join(world_cup_matches, world_cups, by = "Year")

# Print the merged data
print(merged_data)
```

